Spectral Experts for Estimating Mixtures of Linear Regressions

### 2013-06-07 doc/icml-presentation/presentation.tex

    \node<1->[style=box] (gen) at (6,0) {\obj{%
      {\bf Generative Models}
      \begin{itemize}
        \item Hidden Markov Models
        \item Conditional Random Fields
        \item PCFGs
        \item \dots
      \end{itemize}
      }};
    %\node<1->[below=0.5cm of gen-models] (disc) {Discriminative Models};
    \node<2->[style=box,below=1.0cm of gen] (disc) {\obj{%
    {\bf Discriminative Models}
      \begin{itemize}
        \item Max-Entropy Markov Model (Toutanova and Manning, 2000)
        \item Discriminative Parsers (Liang et.\ al, 2006)
        \item Semi-Latent Topic Models (Wang and Mori, 2009)
        \item Easy to include features and tend to be more accurate.
      \end{itemize}
      }};
  \end{tikzpicture}
  \end{centering}

@@

\begin{frame}
  \frametitle{Spectral Methods for Generative Models.}
  \begin{centering}
  \begin{tikzpicture}
    % x, y
    \node<1-2>[style=box] (pos) at (0,0) {\objw{6cm}{%
      \includegraphics[width=\textwidth]{figures/pos.png}\\
      \hfill Part-of-Speech Tagging
      }};
    \node<3->[style=box] (dep) at (0,0) {\objw{6cm}{%
      \includegraphics[width=\textwidth]{figures/dep-parse.png}\\
      \hfill Dependency Parsing
      }};

    %\node<2-> at (6,0) {Discriminative Models};
    \node<1-2>[style=box] (ahk) at (6,0) {\objw{8cm}{%
      \includegraphics[width=\textwidth,height=4cm,keepaspectratio]{figures/three-view.png}\\
      \begin{itemize}
        \item Solution for multi-view case using tensor factorization (Anandkumar et.\ al, 2012).
        \item<2> HMMs (Anandkumar et.\ al, 2012) 
        \item<2> LDA (Anandkumar et.\ al, 2012) 
      \end{itemize}
      \hfill 
      }};
      \node<3->[style=box] at (6,0) {\objw{8cm}{%
      \includegraphics[width=\textwidth,height=4cm,keepaspectratio]{figures/unmixing.png}\\
      \begin{itemize}
        \item Reduce the problem to multi-view learning via tree-unmixing (Hsu et.\ al, 2012). 
      \end{itemize}
      }};
      \node<4->[style=box] at (4,-3) {\objw{12cm}{%
      {\bf How can we apply these methods for discriminative models?}
      }};
  \end{tikzpicture}
  \end{centering}
  \note[item]{\todo{Talk about observable operator stuff.}}
\end{frame}

@@

    \note[item]{I'm unsure whether this stuff really belongs here.}
    \note[item]{We approach the problem by using the method of moments on three conditional views. }
    \note[item]{Consider a moment map $\mathcal{M}$ that maps the parameters $\theta$ to the moments $m$. For a Gaussian, we have that $\mathcal{M} = (\mu, \sigma^2)$.}
    \note[item]{In general, we will compute the inverse of the moment map to learn the parameters from the sample estimates.}
    \note[item]{By the central limit theorem, our sample estimates of the moments converge at a $1/\sqrt{n}$ rate, so we expect that our parameters will also converge at this rate.}
\end{frame}

\begin{frame}
  \frametitle{Method of Moments for Generative LVMs.}

  \begin{centering}
  \begin{tikzpicture}
    % x, y
    \node<1->[style=box]  (moments) at (0,0) {\objw{12cm}{%
      \begin{align*}
        \underbrace{\E[x]}_{M_1} &= \sum_{h=1}^k \pi_h \beta_h & 
        \underbrace{\E[x\tp{2}]}_{M_2} &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & 
        \underbrace{\E[x\tp{3}]}_{M_3} &= \sum_{h=1}^k \pi_h \beta_h\tp{3}. 
      \end{align*}
    }};
    \node<1>[style=box,below=0.1cm of moments] {\objw{12cm}{%
    \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/moments.png}
      }};
    \node<2->[style=box,below=0.1cm of moments] {\objw{12cm}{%
      \begin{itemize}
        \item<2-> {\bf Tensor Power Method} for orthonormal $\beta_h$,
          \begin{align*}
            M_3(\beta_h,\beta_h) 
              &= \sum_{h'=1}^k \pi_{h'} (\beta_{h}^T \beta_{h'})^2 \beta_{h'} \\
              &= \sum_{h'=1}^k \pi_{h'} \delta_{hh'} \beta_{h'} \\
              &= \pi_{h'} \beta_{h}.
          \end{align*}
        \item<3-> Use $M_2$ to whiten $M_3$.
      \end{itemize}
      }};
  \end{tikzpicture}
  \end{centering}

  \begin{centering}
  \begin{tikzpicture}
    % x, y
    \node[style=box] at (0,0) {\objw{36em}{%
        \begin{align*}
          \action<1->{%
            \E[y] &= \E[\beta_h^T x + \epsilon] \\
                  &= M_1^T \E[x]\\
          }
          \action<2->{%
            \E[y x] &= \E[(\beta_h^T x) x + \epsilon x] \\
              &= M_1^T \E[x\tp{2}]\\
          }
          \action<3->{% 
            \E[y^2] &= \E[(\beta_h^T x)^2 + 2 \epsilon (\beta_h^T x) + \epsilon^2 ] \\
                    &= \E[\trace(\beta_h \beta_h^T x x^T)] + \E[\epsilon^2 ] \\
                    &= \trace(M_2 \E[x\tp{2}]) + \E[\epsilon^2 ] \\
          }
          \action<4->{% 
            \E[y^p x\tp{q}] 
                    &= \trace(M_p \E[x\tp{p+q}]) + \cdots.
          }
        \end{align*}
      }};
  \end{tikzpicture}
  \end{centering}
      \note[item]{In the discriminative case, we do not observe $M_2$ directly from the data, but rather, as we'll see, linear measurements of them.}
      \note[item]{Try to work out the moments for this case and see how they {\tt \#fail} epically.}
\end{frame}

>>>>

    \node<1->[style=box] (gen) at (6,0) {\obj{%
      {\bf Generative Models}
      \begin{itemize}
        \item Hidden Markov Models
        \item Conditional Random Fields
        \item PCFGs
        \item \dots
      \end{itemize}
      }};
    %\node<1->[below=0.5cm of gen-models] (disc) {Discriminative Models};
    \node<2->[style=box,below=1.0cm of gen] (disc) {\obj{%
    {\bf Discriminative Models}
      \begin{itemize}
        \item Max-Entropy Markov Model (Toutanova and Manning, 2000)
        \item Discriminative Parsers (Liang et.\ al, 2006)
        \item Semi-Latent Topic Models (Wang and Mori, 2009)
        \item {\bf Easy to include features and tend to be more accurate.}
      \end{itemize}
      }};
  \end{tikzpicture}
  \end{centering}

@@

\begin{frame}
  \frametitle{Spectral Methods}
  \begin{itemize}
    \item Subspace identification [Ljung, 1987]
    \item Observable operator models [Jaeger, 2000]
    \item {\bf Hidden Markov models [Hsu et al, 2009]}
    \item Low-treewidth graphs [Parikh et al., 2012]
    \item Weighted finite state automata [Balle \& Mohri, 2012]
  \end{itemize}
  \begin{itemize}
    \item Mixture of Gaussians [Kalai/Moitra/Valiant, 2010]
    \item {\bf Mixture models, HMMs [Anandkumar/Hsu/Kakade, 2012]}
    \item Latent Dirichlet Allocation [Anandkumar/Hsu/Kakade, 2012]
    \item Stochastic block models [Anandkumar/Ge/Hsu/Kakade, 2012]
    \item Linear Bayesian networks [Anandkumar/Hsu/Javanmard/Kakade, 2012]
  \end{itemize}
  \note[item]{\todo{Talk about observable operator stuff.}}
\end{frame}

\begin{frame}
  \frametitle{Spectral Methods: Tensor Factorization}
  \begin{centering}
  \begin{tikzpicture}
    % x, y
    \node<1>[style=box] (moments) at (0,0) {\objw{12cm}{%
    \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/moments.png}
      }};
    \node<2->[style=box] (moments) at (0,0) {\objw{12cm}{%
      \begin{itemize}
        \item<2-> {\bf Tensor Power Method} for orthonormal $\beta_h$,
          \begin{align*}
            M_3(\beta_h,\beta_h) 
              &= \sum_{h'=1}^k \pi_{h'} (\beta_{h}^T \beta_{h'})^2 \beta_{h'} \\
              &= \sum_{h'=1}^k \pi_{h'} \delta_{hh'} \beta_{h'} \\
              &= \pi_{h'} \beta_{h}.
          \end{align*}
        \item<3-> Use $M_2$ to whiten $M_3$.
      \end{itemize}
      }};
    \node<1->[style=box,below=0.1cm of moments]  {\objw{12cm}{%
      \begin{align*}
        \underbrace{\E[x]}_{M_1} &= \sum_{h=1}^k \pi_h \beta_h & 
        \underbrace{\E[x\tp{2}]}_{M_2} &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & 
        \underbrace{\E[x\tp{3}]}_{M_3} &= \sum_{h=1}^k \pi_h \beta_h\tp{3}. 
      \end{align*}
    }};
  \end{tikzpicture}
  \end{centering}

\begin{comment}
\begin{frame}
  \frametitle{Spectral Methods for Generative Models.}
  \begin{centering}
  \begin{tikzpicture}
    % x, y
    \node<1-2>[style=box] (pos) at (0,0) {\objw{6cm}{%
      \includegraphics[width=\textwidth]{figures/pos.png}\\
      \hfill Part-of-Speech Tagging
      }};
    \node<3->[style=box] (dep) at (0,0) {\objw{6cm}{%
      \includegraphics[width=\textwidth]{figures/dep-parse.png}\\
      \hfill Dependency Parsing
      }};

    %\node<2-> at (6,0) {Discriminative Models};
    \node<1-2>[style=box] (ahk) at (6,0) {\objw{8cm}{%
      \includegraphics[width=\textwidth,height=4cm,keepaspectratio]{figures/three-view.png}\\
      \begin{itemize}
        \item Solution for multi-view case using tensor factorization (Anandkumar et.\ al, 2012).
        \item<2> HMMs (Anandkumar et.\ al, 2012) 
        \item<2> LDA (Anandkumar et.\ al, 2012) 
      \end{itemize}
      \hfill 
      }};
      \node<3->[style=box] at (6,0) {\objw{8cm}{%
      \includegraphics[width=\textwidth,height=4cm,keepaspectratio]{figures/unmixing.png}\\
      \begin{itemize}
        \item Reduce the problem to multi-view learning via tree-unmixing (Hsu et.\ al, 2012). 
      \end{itemize}
      }};
      \node<4->[style=box] at (4,-3) {\objw{12cm}{%
      {\bf How can we apply these methods for discriminative models?}
      }};
  \end{tikzpicture}
  \end{centering}
  \note[item]{\todo{Talk about observable operator stuff.}}
\end{frame}
\end{comment}

@@

    \note[item]{I'm unsure whether this stuff really belongs here.}
    \note[item]{We approach the problem by using the method of moments on three conditional views. }
    \note[item]{Consider a moment map $\mathcal{M}$ that maps the parameters $\theta$ to the moments $m$. For a Gaussian, we have that $\mathcal{M} = (\mu, \sigma^2)$.}
    \note[item]{In general, we will compute the inverse of the moment map to learn the parameters from the sample estimates.}
    \note[item]{By the central limit theorem, our sample estimates of the moments converge at a $1/\sqrt{n}$ rate, so we expect that our parameters will also converge at this rate.}
\end{frame}
\end{comment}

% \begin{frame}
%   \frametitle{Method of Moments for Generative LVMs.}
% 
%   \begin{centering}
%   \begin{tikzpicture}
%     % x, y
%     \node<1->[style=box]  (moments) at (0,0) {\objw{12cm}{%
%       \begin{align*}
%         \underbrace{\E[x]}_{M_1} &= \sum_{h=1}^k \pi_h \beta_h & 
%         \underbrace{\E[x\tp{2}]}_{M_2} &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & 
%         \underbrace{\E[x\tp{3}]}_{M_3} &= \sum_{h=1}^k \pi_h \beta_h\tp{3}. 
%       \end{align*}
%     }};
%     \node<1>[style=box,below=0.1cm of moments] {\objw{12cm}{%
%     \includegraphics[width=\textwidth,height=6cm,keepaspectratio]{figures/moments.png}
%       }};
%     \node<2->[style=box,below=0.1cm of moments] {\objw{12cm}{%
%       \begin{itemize}
%         \item<2-> {\bf Tensor Power Method} for orthonormal $\beta_h$,
%           \begin{align*}
%             M_3(\beta_h,\beta_h) 
%               &= \sum_{h'=1}^k \pi_{h'} (\beta_{h}^T \beta_{h'})^2 \beta_{h'} \\
%               &= \sum_{h'=1}^k \pi_{h'} \delta_{hh'} \beta_{h'} \\
%               &= \pi_{h'} \beta_{h}.
%           \end{align*}
%         \item<3-> Use $M_2$ to whiten $M_3$.
%       \end{itemize}
%       }};
%   \end{tikzpicture}
%   \end{centering}
% 
%     \note[item]{\bf We can recover the means of a generative LVM by exploiting their symmetric structure.}
%     \note[item]{\begin{align*} M_2 &= \sum_{h=1}^k \pi_h \beta_h\tp{2} & M_3 &= \sum_{h=1}^k \pi_h \beta_h\tp{3}.  \end{align*}}
%       \note[item]{In general, $M_2$ is insufficient to identify the model, because it is invariant to rotations of $B$. }
%       \note[item]{If the $\beta_h$ were orthogonal, then the eigenvectors and eigenvalues of $M_3$ would give us the answer. }
%       \note[item]{In general, we need both $M_2$ and $M_3$. We can use the whitening transformation for $M_2$ to whiten $M_3$ such that it has a orthogonal decomposition.}
%       \note[item]{The robust tensor power method by \cite{AnandkumarHsuGe2012} find stable eigenvectors.}
% \end{frame}

\begin{comment}
\begin{frame}
  \frametitle{Method of Moments for the Mixture of Linear Regressions.}

  \begin{centering}
  \begin{tikzpicture}
    % x, y
    \node[style=box] at (0,0) {\objw{36em}{%
        \begin{align*}
          \action<1->{%
            \E[y] &= \E[\beta_h^T x + \epsilon] \\
                  &= M_1^T \E[x]\\
          }
          \action<2->{%
            \E[y x] &= \E[(\beta_h^T x) x + \epsilon x] \\
              &= M_1^T \E[x\tp{2}]\\
          }
          \action<3->{% 
            \E[y^2] &= \E[(\beta_h^T x)^2 + 2 \epsilon (\beta_h^T x) + \epsilon^2 ] \\
                    &= \E[\trace(\beta_h \beta_h^T x x^T)] + \E[\epsilon^2 ] \\
                    &= \trace(M_2 \E[x\tp{2}]) + \E[\epsilon^2 ] \\
          }
          \action<4->{% 
            \E[y^p x\tp{q}] 
                    &= \trace(M_p \E[x\tp{p+q}]) + \cdots.
          }
        \end{align*}
      }};
  \end{tikzpicture}
  \end{centering}
      \note[item]{In the discriminative case, we do not observe $M_2$ directly from the data, but rather, as we'll see, linear measurements of them.}
      \note[item]{Try to work out the moments for this case and see how they {\tt \#fail} epically.}
\end{frame}
\end{comment}
